---
title: "Getting Started with authClient"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with authClient}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

This walks through wiring `authClient` into a Shiny app from scratch, then
extending that to real single sign-on (SSO) across two separate apps
sharing the same identity database.

Code chunks below are shown for reading, not executed as part of building
this vignette — a real Shiny app blocks, which doesn't fit inside a knitted
document. Copy them into an actual `app.R` to run them.

## 1. What you need before starting

- A shared identity database already following the schema in
  `python/repository/tablestructure_user.py` (`users`, `apps`,
  `user_app_access`, `roles`, `permissions`, `user_roles`, `sessions`, etc.),
  reachable from your app's process.
- An `app_key` registered for your app in the `apps` table — this is
  normally created by a console admin through `accessConsole`'s "Create a
  new app" panel, not by your app itself. `authClient` only *reads* access
  and permission grants; it doesn't manage them.
- An `AUTH_SESSION_SECRET` environment variable, set to the **same** value
  in every app that should share SSO sessions. This is what lets one app
  validate a session cookie another app issued — different secrets means
  the tokens simply won't validate anywhere else.

## 2. A single app, with login and an access gate

```{r}
library(shiny)
library(pool)
library(authClient)

con <- dbPool(RSQLite::SQLite(), dbname = "path/to/shared.sqlite", onCreate = function(conn) {
  DBI::dbExecute(conn, "PRAGMA foreign_keys = ON;")
  DBI::dbExecute(conn, "PRAGMA journal_mode = WAL;")
})
ensure_sessions_schema(con)
ensure_auth_settings_schema(con)
onStop(function() poolClose(con))

ui <- fluidPage(
  authLoginUI("auth"),
  uiOutput("app_body")
)

server <- function(input, output, session) {
  auth <- authLoginServer("auth", con, require_app_key = "invoicing")

  output$app_body <- renderUI({
    # checked() only flips to TRUE once the SSO cookie check has actually
    # resolved -- gating on logged_in() alone flashes the login form for a
    # frame on every page load.
    req(auth$checked())
    if (!auth$logged_in()) return(NULL)

    tagList(
      fluidRow(
        column(9, h4(paste("Logged in as", auth$username()))),
        column(3, actionButton("logout_btn", "Log out"))
      ),
      p("You're in -- user_has_app_access() already passed, or you wouldn't be here.")
    )
  })

  observeEvent(input$logout_btn, auth$logout())
}

shinyApp(ui, server)
```

If the logged-in user doesn't have access to `"invoicing"` (checked via
`user_has_app_access()` internally), the login form shows "You don't have
access to this application." instead of letting them in — access is denied
at login time, not left for the rest of the app to check.

## 3. Gating a specific tab on a permission, not just app access

`has_permission()` is checked separately from app access — a user can be
allowed to *use* an app but not hold every permission inside it (e.g. can
view invoices but not delete them):

```{r}
output$app_body <- renderUI({
  req(auth$checked())
  if (!auth$logged_in()) return(NULL)

  tagList(
    tabsetPanel(
      tabPanel("View", p("Everyone with app access sees this.")),
      if (auth$has_permission("invoicing", "invoice.delete")) {
        tabPanel("Delete", actionButton("del", "Delete an invoice"))
      }
    )
  )
})
```

`has_permission()` results are cached for up to 8 hours per session by
default (`permission_cache_ttl_secs` on `authLoginServer()`) — fine for
deciding whether a tab is visible, not a substitute for checking again at
the point a sensitive action actually runs.

## 4. Letting users change their own password

```{r}
output$app_body <- renderUI({
  req(auth$checked())
  if (!auth$logged_in()) return(NULL)
  changePasswordUI("pw")
})
changePasswordServer("pw", con, auth$user_id)
```

A successful change revokes every other session belonging to that user —
an old, still-open browser tab elsewhere won't keep working on the old
password.

## 5. Turning this into real SSO across two apps

Nothing extra to *build* — SSO is just what you get for free once two apps
satisfy the requirements from step 1:

1. Both apps point at the **same** database.
2. Both apps have the **same** `AUTH_SESSION_SECRET`.
3. Both apps are served under a **shared parent domain** (e.g.
   `invoicing.internal.example.com` and `reporting.internal.example.com`),
   so a cookie scoped with `cookie_domain = ".internal.example.com"` reaches
   both. Pass that explicitly to `authLoginServer()` on every app:

```{r}
auth <- authLoginServer("auth", con,
                         require_app_key = "reporting",
                         cookie_domain = ".internal.example.com")
```

With that in place: log into the invoicing app, then open the reporting
app in the same browser — its own `authLoginServer()` reads the same
cookie on page load, calls `validate_session()`, and logs you in silently.
No second password prompt, and (if audit logging is enabled — see below)
both apps show up as separate entries in the audit log, because the
cookie-reuse path logs access just like a fresh login does.

If your apps *aren't* under a shared parent domain, the cookie's `Domain=`
scope can't reach across them — you'd need a central auth gateway in front
of everything instead, which is a much bigger infrastructure project than
this package covers.

## 6. Turning on presence auditing (optional, off by default)

```{r}
set_audit_log_enabled(con, TRUE)
```

This is one flag in the shared `auth_settings` table — flipping it applies
to every app using `authClient` against this database immediately, no
per-app config or redeploy. It only ever records *that* a user accessed a
given app (`user_id`, `app_id`, timestamp) — not what they did or for how
long.
